---
title: "Assignment 1"
author: Emma
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: kate
    smart: yes
    theme: cosmo
    code_folding: show
    df_print: paged
editor_options: 
  chunk_output_type: inline
---

**Due Date**: Sunday, March 24th

```{r, message=FALSE}
library(tidyverse) ; library(nycflights13) ; library(dplyr)
```
Day 1
#1Get the structure of the flights data set using the str() command.


Use the pull() function instead of the $ notation to achieve the same result as above. Save this as pulledCarrier
```{r}
flights %>% 
  pulledCarrier <- pull(carrier)
```


Use this function identical() to establish that the pulledCarrier and flightCarrier are the same.


Now try the same step as 8 with select() instead of pull. Save this as selectedCarrier


Use the identical() function to establish that selectedCarrier and flightCarrier are not the same.


Why aren’t selectedCarrier and flightCarrier identical? Think about the data structure that select() returns compared to that for pull()/$
