---
title: "Assignment 1"
author: Emma
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: kate
    smart: yes
    theme: cosmo
    code_folding: show
    df_print: paged
editor_options: 
  chunk_output_type: inline
---

**Due Date**: Sunday, March 24th

```{r, message=FALSE}
library(tidyverse) ; library(nycflights13) ; library(dplyr)
```
Day 1
#1Get the structure of the flights data set using the str() command.

```{r}
str(flights)
```

#The $ notation is used in base R to pull out columns from a data frame. Its usage is as follows df$columnName, where df is the dataframe (or tibble) you are interested in pulling a column denoted by columnName from. Use this syntax to pull out the carrier column from the flights tibble and store it in a variable called flightCarrier.

```{r}
flightCarrier <- flights$carrier

```


#3Get the length of the flightCarrier vector and compare it with the number of rows in the flights data using the == operator. It should return true. Search the internet for the function to find the number of rows of a dataframe.

```{r}
flights %>% 
  length(carrier) == length(flightCarrier)
```


#4Use the unique() function on the flightCarrier variable you created in the previous step to find the unique list of carriers in the flights dataset. Replace the flightCarrier with this new vector of unique values.

```{r}

 flightCarrier <- unique(flightCarrier)

```


#5Print the new length of the flightCarrier vector.

```{r}
length(flightCarrier)
```


#6Get the data type of the flightCarrier vector using typeof(), is it the same as the type denoted in str(). (The output from the str() function abbreviates the type so for instance integer is denoted as int.)
```{r}
typeof(flightCarrier)

str(flightCarrier)
```


#7Use pipes to combine steps 2 and 4. Like so df$columnName %>% unique().
```{r}

```


Use the pull() function instead of the $ notation to achieve the same result as above. Save this as pulledCarrier
```{r}
flights %>% 
  pulledCarrier <- pull(carri)
```


Use this function identical() to establish that the pulledCarrier and flightCarrier are the same.


Now try the same step as 8 with select() instead of pull. Save this as selectedCarrier


Use the identical() function to establish that selectedCarrier and flightCarrier are not the same.


Why aren’t selectedCarrier and flightCarrier identical? Think about the data structure that select() returns compared to that for pull()/$
