---
title: "Class 13 - In class exercises"
author: Hari Subhash
date: "Date created: `r Sys.Date()`"
output:
  html_notebook:
    highlight: kate
    smart: yes
    theme: cosmo
    code_folding: show
    df_print: paged
---
##Task Set 1: Repeat after me
1. The `table1` below is a tidy dataset. `gather()` the columns case and population to create an untidy table that mirrors `table2`. See question 2 below to see how `table2` looks.
#Gather and spread mirror each other in terms of syntax. Gather gathers columbs into key value pairs. Spread spreads key value pairs into new columns. 
#Separate separates a combined column into multiple. Combine combines multiple columns into a single column.

```{r}
table1
```


```{r, message=FALSE}
library(tidyverse) ; 
table1 %>% 
     gather(key = "type", value = "count", cases , population)
```


```{r}
table2
```

2. Use `spread()` to spread the `type` column into a wide format.
```{r}
unique(table2$type)
```

```{r}
table2 %>% 
  spread(key = type, value = count)
```


```{r}
table3 %>% 
  separate(col = rate, into = c("cases" , "population") , sep = "/")
```

3. `table4a` contains the values for the count variable.

```{r}
table4a
```
Here is how I would clean this.
```{r}
table4a %>% 
    gather(key = "year", value = "value", -country)
```

#another way

```{r}
table4a %>% 
    gather(key = "year", value = "value", `1999`, `2000`)
```

Use a similar approach to clean `table4b` and join it (using a `left_join()`) with the cleaned `table4a` to recreate table1. Bonus points if you do this without storing either table4b or table4a in a separate variable.
```{r}
table4a %>% 
    gather(key = "year", value = "value", `1999`, `2000`) %>% 
left_join(., table4b %>% 
    gather(key = "year", value = "value", `1999`, `2000`), 
  by = c("country", "year")) %>% 
  arrange(country, year)
  
```

```{r}

```

##Task Set 2: Identify issues
1. Load the messy gapminder data using `read_csv()`
2. Discuss the issues with this data within your group (from the perspective of tidyness)
3. Write down the steps you will take to clean this data up (in terms of `tidyr` functions). So for instance: "I will gather xyz columns, then separate a_b_c...."

##Task Set 3: Tidy
1. Load and clean the untidy gapminder data.

