---
title: "R Notebook"
output: html_notebook
---

##List the 4 most common (data) types of atomic vectors.
##Answer: 

##2.Create atomic vectors of the 4 common types covered in class. Print their length and type.


##3. List the heirarchy of data types for coercion
##Answer:

##4. Explain why the output of following bits of code is TRUE (explain what is happening in terms of coercion):
```{r}
flights %>%
  top_n(., n=5, wt = dep_delay)
```
##1.Load the tidyverse and nycflights13 packages.
```{r, message=FALSE}
library(nycflights13) ; library(tidyverse)
```

##.2 Describe two ways in which you can use RStudio confirm that a package has been loaded into your current R session? (Hint: One of them is a dropdown menu and the other a list of checkboxes)
```{r}

```

##Answer: You can check the console, where it states that [Workplace loaded from ~?.RData], and 

##3.Use the distinct() command to identify the unique combinations of airline carriers and origin airports in the flights data.
```{r}
flights %>%
  distinct(.,carrier, origin, keep_all=TRUE)
```

```{r}
unique(flights$carrier)
```


Use the one_of() helper function to select the following columns - year, month, day, dep_time, sched_dep_time, dep_delay, arr_time, from the flights data.

```{r}

```


##Use the slice() function to slice the first 10 rows in the flights data.

```{r}

```


##Perform the same operation as the previous task but this time use filter(). (Hint: Search for how to get the row numbers of a tibble)


##Use the top_n() command to identify the top 5 flights out of NYC that experienced the greatest arrival delays in 2013.

```{r}
flights %>% 
  top_n(., n = 5, wt = arr_delay) %>% 
  arrange(desc(arr_delay))

```


##Use top_n() along with arrange() and slice() to pull out the 50th most delayed departure from NYC in 2013 (Not all 50, just the 50th).

```{r}
flights %>% 
  top_n(., n = 50, wt = dep_delay) %>% 
  arrange(desc(dep_delay)) %>% 
  slice(50)
```


##Do the same thing as the previous task using
row_number()
and dense_rank()

```{r}
flights %>% 
 filter(row_number(desc(dep_delay)) == 50)

```

```{r}
flights %>% 
filter(dense_rank(desc(dep_delay)) == 50)
```


##The result from row_number() and top_n() are the same, while that from dense_rank() is different. Why?



Answer:
##Find the top 10 airline carrier that had the highest average departure delays in 2013 using group_by(), summarise() and other functions you have learnt previously.
```{r}
flights %>% 
  group_by(carrier) %>% 
  summarise(mean(dep_delay), keep_all )
```

##Use group_by() with mutate() to create a new variable called comparativeDelay which is the difference between departure delay and the average delay in each origin airport for every hour in 2013 (checkout the time_hour variable in the flights data). Store the result in a variable called comparativeDelays.

```{r}
flights %>% 
  group_by(origin) %>% 
  mutate(., x = origin , y =  ) %>% 
  varName <- comparativeDelay

```


##Now group the comparativeDelays tibble by carriers to print the top 10 airlines with the greatest average comparative delays.



##Use group_by() with filter to print the 5 most delayed flights from each origin. Your printed tibble should have 15 rows.



##The air authority in NY wants to penalize carriers for departure delays. Which of the three metrics (task 1, 3 or 4) would you recommend they use to identify the airlines to penalize. Why?

Answer: