---
title: "Week 3 - Daily R Nuggets"
author: "Emma L"
date: 'Date created: `r Sys.Date()`'
output:
  html_notebook:
    code_folding: show
    df_print: paged
    highlight: kate
    smart: yes
    theme: cosmo
  html_document:
    df_print: paged
---


```{r, message = FALSE}
library(nycflights13) ; library(tidyverse)
```

##Day 1
1. Get the structure of the flights data set using the `str()` command.
```{r}
  str(flights)
```



2. The `$` notation is used in base R to pull out columns from a data frame. Its usage is as follows `df$columnName`, where df is the dataframe (or tibble) you are interested in pulling a column denoted by columnName from. Use this syntax to pull out the `carrier` column from the flights tibble and store it in a variable called `flightCarrier`.
```{r}
flightCarrier <- flights$carrier
```

3. Get the length of the `flightCarrier` vector and compare it with the number of rows in the flights data using the `==` operator. It should return true. Search the internet for the function to find the number of rows of a dataframe.
```{r}
length(flightCarrier)

nrow(flights)

length(flightCarrier) == nrow(flights)
```
4. Use the `unique()` function on the `flightCarrier` variable you created in the previous step to find the unique list of carriers in the flights dataset. Replace the `flightCarrier` with this new vector of unique values.

```{r}
flightCarrier <- unique(flightCarrier)
```

5. Print the new length of the `flightCarrier` vector.
```{r}
length(flightCarrier)
```

6. Get the data type of the `flightCarrier` vector using `typeof()`, is it the same as the type denoted in `str()`. (The output from the `str()` function abbreviates the type so for instance integer is denoted as int.)
```{r}
typeof(flightCarrier)

str(flightCarrier)
```

7. Use pipes to combine steps 2 and 4. Like so `df$columnName %>% unique()`.
```{r}
flightCarrier <- flights$carrier %>% unique()

#I tested this using the length command below, but why does it not work if you put the new var "flightCarrier into the unique bracket?"   

length(flightCarrier)
str(flightCarrier)

```

8. Use the `pull()` function instead of the `$` notation to achieve the same result as above. Save this as `pulledCarrier`
```{r}
pulledCarrier <- flights %>% pull(carrier) %>% unique()

length(pulledCarrier)
nrow(pulledCarrier)
ncol(pulledCarrier)
str(pulledCarrier)
```

9. Use this function `identical()` to establish that the `pulledCarrier` and `flightCarrier` are the same.
```{r}
identical(pulledCarrier, flightCarrier)
```

10. Now try the same step as 8 with `select()` instead of pull. Save this as `selectedCarrier`
```{r}

selectedCarrier <- flights %>% select(carrier) %>% unique()

length(selectedCarrier)
nrow(selectedCarrier)  
str(selectedCarrier)
```

11. Use the `identical()` function to establish that `selectedCarrier` and `flightCarrier` are not the same. 
```{r}
identical(selectedCarrier, flightCarrier)
```

12. Why aren't `selectedCarrier` and `flightCarrier` identical? Think about the data structure that `select()` returns compared to that for `pull()`/`$`

#Answer: The *select()* function takes a column as 1 variable (with 16 observations, using the *unique() function). So its length is 1 (R counts it as 1 variable) but its number of unique rows is 16
#Whereas the *pull()* function extracts only the column **values** as a vector. (It does not pull the column name, so the *str()* command shows only that it is a character vector, and does not show that the values are contained within a data frame.)

##Day 2
1. Arrange flights in the descending order of arrival delay and ascending order of departure delay.

```{r}
  arrange(flights, desc(arr_delay))
```

```{r}
arrange(flights, dep_delay)
```

2. Use three different ways of your choice to select the first 5 columns from the flights tibble. (using number . 

```{r}
select(flights, "year" , "month" , "day" , "dep_time", "sched_dep_time")
```

```{r}
firstFive <- c("year" , "month" , "day" , "dep_time", "sched_dep_time")
  select(flights, firstFive)
```


3. Which airports had the most number of flights that were delayed by more than 10 minute in 2013? (You can show a table with the counts of flights delayed by more than 10 minutes from each origin that is sorted in descending order of the count)

```{r}
flights %>% 
  group_by(origin) %>% 
  filter(dep_delay > 10) %>% 
  tally(dep_delay, sort = T)

```
#My own question to inform the last answer: Which airline had the most number of flights delayed?
```{r}
flights %>% 
  group_by(carrier) %>% 
  filter(dep_delay > 10) %>%
  tally(dep_delay, sort = T)
  
  
 
```

4. Create a new variable `timeLostInAir` that is the difference between arrival delay and departure delay. Save this to a variable.

```{r}
flights %>% 
  mutate(timeLostInAir = dep_delay - arr_delay) %>% 
  select(carrier, 6:10, origin, timeLostInAir) %>% 
  arrange(desc(timeLostInAir))

```

5. Use the tibble saved in task 4 to find the flight that lost the most amount of time while in air
```{r}
flights %>% 
  mutate(timeLostInAir = dep_delay - arr_delay) %>% 
  arrange(desc(timeLostInAir)) %>% 
  top_n(timeLostInAir, n=1) %>% 
  select(carrier, origin, dest, timeLostInAir)
```

6. Use the tibble saved in task 4 to show the top 5 carriers in the terms of `timeLostInAir`
```{r}
flights %>% 
  group_by(carrier) %>% 
  mutate(timeLostInAir = dep_delay - arr_delay) %>%
  summarise(timeLostInAir = sum(timeLostInAir, na.rm = T)) %>% 
  ungroup() %>% 
  
  top_n(n = 5, timeLostInAir) %>% 
  arrange(desc(timeLostInAir))
 
    
```


7. Do you think that `timeLostInAir` is a useful variable to continue exploring based on the last 3 tasks?

#Yes, but I would also want to use it to see what other variables might have a strong correlation to timeLostinAir (EG distance, departure time). 

##Day 3
1. Convert the `dest` variable in the flights tibble to a factor using the `mutate()` and pipe it to an `str()` function to see that the `dest` variable is now marked as a factor. (The task is to change the `dest` variable, not add a new column to the tibble)
```{r}
flights %>%
  mutate(dest = factor(dest)) %>% 

str(flights)
```

2. Use `mutate()` along with `recode()` to recode the origin column as follows: JFK to John F Kennedy, LGA to La Guardia, leave EWR as is. Read the documentation for recode on tidyverse to understand its implementation. 
```{r}
origin <- flights %>% 
  mutate(recode(origin, "JFK" = "John F. Kennedy" , "LGA" = "LaGuardia")) %>% 
  
str(flights)
```

3. Repeat task 2 but this time use `mutate()` along with `case_when()`. If you have further questions post on slack.
```{r}
flights %>% 
  mutate(origin, case_when(origin == "JFK" ~ "John F. Kennedy",
                   origin == "LGA" ~ "LaGuardia",
                   origin == "EWK" ~ "EWK"
                   )) %>% 
 
str(flights)
```


4. Based on the flights 2013 data, which are the worst 5 destinations in terms of arrival delay?
```{r}
flights %>%
  group_by(dest) %>%
  summarise(meanDelay = mean(arr_delay, na.rm = T)) %>% 
  arrange(desc(meanDelay)) %>% 
  ungroup() %>% 
  
  slice(1:5)
```
#Columbia, SC, Tulsa, OK, Oklahoma City, OK, Jackson Hole, WY, Knoxville Tyson Airport
