---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse) ; library(nycflights13)
```

```{r}
library(lubridate)
```

##1. Select the first three columns (year, month, day), all other columns whose names `contains()` the term "delay" and the origin column. Filter this data to show all the flights that took off in the morning (before 12:00) from JFK in December. Make sure to use pipes between the select and filter command. Refer to the shortcut for inserting a pipe (see above).
```{r}
flights %>%
  select(1:3, contains("dep"), origin)%>%
  filter(dep_time < 1200 & origin == "JFK" & month == 12)
```
##2. Use `top_n()` to print the most delayed departures from NYC in 2013. Read the documentation for `top_n()` on tidyverse if you are confused.

```{r}
flights %>%
  top_n(., n=5, wt = dep_delay)
```

##1. Create a variable that indicates whether a flight took off in the AM or the PM.

```{r}
paste("R4D", "class" , "is" , "awesome" , sep = "-")
```
```{r}
flightsWeek <- flights %>%
  mutate(amFlight = if_else(sched_dep_time < 1200 , "morning" , "evening"))
```


##2. Use `transmute()` instead of `mutate()` to do the same. What is the difference between the two?

```{r}
flights %>%
  transmute(amFlight = if_else(sched_dep_time < 1200, "morning" , "evening"))
```




##Grouped operations (Task time 30 mins)
##1. Find the top 10 airline carrier that had the highest average departure delays in 2013 using `group_by()`, `summarise()` and other functions you have learnt previously.

```{r}
flights %>%
  group_by(origin) %>%
  summarise((avgDelay = mean(dep_delay, na.rm = T)))
```

```{r}
flights %>%
  group_by(origin) %>%
  summarise((avgDelay = mean(dep_delay, na.rm = T)))
```


##2. Use `group_by()` with `mutate()` to create a new variable called `comparativeDelay` which is the difference between departure delay and the average delay in each origin airport for every hour in 2013 (checkout the `time_hour` variable in the flights data). Store the result in a variable called `comparativeDelays`.
##3. Now group the `comparativeDelays` tibble by carriers to print the top 10 airlines with the greatest average comparative delays.
##4. Use `group_by()` with filter to print the 5 most delayed flights from each origin. Your printed tibble should have 15 rows.
5. The air authority in NY wants to penalize carriers for departure delays. Which of the three metrics (task 1, 3 or 4) would you recommend they use to identify the airlines to penalize. Why?


